Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-15 aistats.tex
pass on optimization section

As we saw in \S\ref{sec:problem}, MAP can be interpreted as stochastic mirror descent (SMD).
This means that \textbf{1)} we may obtain a convergence rate for MAP from an optimization analysis, and \textbf{2)} any insights gained from the MAP may inform further design and analysis of SMD.
We first review the assumptions of relative smoothness, helpful to deal with non-smooth functions, before investigating recent analyses of SMD with the MAP.

@@

To analyze stochastic algorithms, one also need to quantify the randomness of stochastic gradients $g(\nat)$ .
While many assumptions exist for SGD \citep[\S3 for a modern review]{khaled2020better}, only a few have been adapted to SMD with relative smoothness \citep{hanzely2018fastest, dragomir2021fast, dorazio2021stochastic}, but they have so far been lacking concrete examples.
We review these analyses in the light of the MAP and provide a summary in \cref{tbl:assumptions}.

@@

\subsubsection{Analogs of the Variance}
let us introduce $\cS_\conj(\m_1, \m_2) = \bregmanconj(\m_1, \m_2) + \bregmanconj(\m_2, \m_1)$ th symmetrized Bregman induced by $\conj$.
\citet{hanzely2018fastest} assume that the expectation of $\cS$ between stochastic and deterministic updates verifies
\alignn{
\expect[g]{\cS_\conj \paren{
	\hat\m_{n} - \lr g(\hat \nat_{n}),
	\hat\m_{n}   -\lr \nabla f(\hat \nat_{n})
}} \leq \lr^2 C
\label{eq:hanzely}
}
for all possible iterates $\MAPt$, relevant step-sizes $\lr$ and for some constant $C$.
When $A(\theta) = \frac{1}{2}\norm{\theta}^2$ or when $A$ is $\stgcvx$-strongly convex,
this definition recovers the variance of the stochastic gradient
\[
	\expect[g\!\!]{\inv{\stgcvx}\norm{\nabla f(\theta) - g(\theta)}^2}\leq C.
\]
Under this assumption, \citet[Lem.4.8]{hanzely2018fastest} prove a $O(1/n)$ convergence rate on function values with $O(1/n)$ step-sizes and tail averaging \citep{lacostejulien2012simpler} in primal space $\Theta$.

@@

However, both of their proofs hold if we add an expectation over $\MAPm$ to their assumption.
But this is not helpful for us, as verifying the assumption become as hard as the initial problem.
For instance, the expectation of~\eqref{eq:hanzely} over $\MAPm$ is an upper-bound on the variance term of~\eqref{eq:primal_pivot}.
Confronted to this difficulty, let us move on to an alternative definition of variance.

>>>>

As we saw in \S\ref{sec:problem}, MAP can be interpreted as stochastic mirror descent (SMD).
This means that \textbf{1)} we may obtain a convergence rate for MAP from an optimization analysis, and \textbf{2)} any insights gained from the MAP may inform other designs and analyses of SMD.
We first review the assumptions of relative smoothness, helpful to deal with non-smooth functions, before investigating recent analyses of SMD with the MAP.

@@

To analyze stochastic algorithms, one also needs to quantify the randomness of stochastic gradients $g(\nat)$.
While many assumptions exist for SGD \citep[\S3 for a modern review]{khaled2020better}, only a few have been adapted to SMD with relative smoothness \citep{hanzely2018fastest, dragomir2021fast, dorazio2021stochastic}, but they have so far been lacking concrete examples.
We review these analyses in the light of the MAP and provide a summary in \cref{tbl:assumptions}.

@@

\subsubsection{Analogs of the Variance}
Let us introduce the symmetrized Bregman induced by $\conj$, written $\cS_\conj(\m_1, \m_2) = \bregmanconj(\m_1, \m_2) + \bregmanconj(\m_2, \m_1)$.
\citet{hanzely2018fastest} assume that the expectation of $\cS$ between stochastic and deterministic updates verifies
\alignn{
	\expect[g]{\cS_\conj \paren{
			\hat\m_{n} - \lr g(\hat \nat_{n}),
			\hat\m_{n}   -\lr \nabla f(\hat \nat_{n})
	}} \leq \lr^2 C
	\label{eq:hanzely}
}
for all possible iterates $\MAPt$, relevant step-sizes $\lr$ and for some constant $C$.
When $A(\theta) = \frac{1}{2}\norm{\theta}^2$ or when $A$ is $\stgcvx$-strongly convex,
this definition recovers the variance of the stochastic gradient
\[
\expect[g\!\!]{\inv{\stgcvx}\norm{\nabla f(\theta) - g(\theta)}^2}\leq C.
\]
Under this assumption, \citet[Lem.4.8]{hanzely2018fastest} prove a $O(1/n)$ convergence rate on function values with $O(1/n)$ step-sizes and tail averaging \citep{lacostejulien2012simpler} in primal space $\Theta$.

@@

However, both of their proofs hold if we add an expectation over $\MAPm$ to their assumption.
However, this is not helpful, as verifying the assumption becomes as hard as the initial problem.
For instance, the expectation of~\eqref{eq:hanzely} over $\MAPm$ is an upper bound on the variance term of~\eqref{eq:primal_pivot}.
Confronted with this difficulty, we investigate an alternative definition of variance.
